---
title: "wk04_tidyr"
author: "Hartline"
date: "February 5, 2016"
output: html_document
---

## joining data

For this portion of the individual assignment, similar to last week, you'll find it easiest to copy and paste from `## joining data` onwards in  [`env-info/wk05_joining.Rmd`](https://raw.githubusercontent.com/ucsb-bren/env-info/gh-pages/wk05_joining.Rmd) to your `env-info_hw/wk05_ggplot.Rmd`. Then you can play with different chunks of the code. Please be sure to answer all tasks and questions at the bottom.



The R chunks explaining the `dplyr` join functions below are pulled from the excellent [**wrangling-webinar.pdf**](wk03_dplyr/wrangling-webinar.pdf) presentation, which you should consult as you execute (see shortcuts in [rstudio-IDE-cheatsheet.pdf](refs/cheatsheets/rstudio-IDE-cheatsheet.pdf)). 

### setup

Ensure that you're in the same working directory `env-info_hw` when you Knit HTML as when you test code in the Console.

```{r setwd students}
wd = 'env-info_hw'

# set working directory for Console (vs Rmd)
if (wd %in% list.files() & interactive()){
    setwd(wd)
}

# ensure working directory
if (basename(getwd()) != wd){
  stop(sprintf("WHOAH! Your working directory is not in '%s'!\n   getwd(): %s", wd, getwd()))
}
```

### `bind_cols`

```{r bind_cols}
library(dplyr)

y = data.frame(
  x1 = c('A','B','C'), 
  x2 = c( 1 , 2 , 3), 
  stringsAsFactors=F)
z = data.frame(
  x1 = c('B','C','D'), 
  x2 = c( 2 , 3 , 4), 
  stringsAsFactors=F)
y
z
bind_cols(y, z)
```

### `bind_rows`

```{r bind_rows}
y
z
bind_rows(y, z)
```

### `union`

```{r union}
y
z
union(y, z)
```

### `intersect`

```{r intersect}
y
z
intersect(y, z)
```

### `setdiff`

```{r setdiff}
y
z
setdiff(y, z)
```

### `left_join`

```{r left_join}
songs = data.frame(
  song = c('Across the Universe','Come Together', 'Hello, Goodbye', 'Peggy Sue'),
  name = c('John','John','Paul','Buddy'), 
  stringsAsFactors=F)
artists = data.frame(
  name = c('George','John','Paul','Ringo'),
  plays = c('sitar','guitar','bass','drums'), 
  stringsAsFactors=F)
left_join(songs, artists, by='name')
```

### `inner_join`

```{r inner_join}
inner_join(songs, artists, by = "name")
```

### `semi_join`

```{r semi_join}
semi_join(songs, artists, by = "name")
```

### `anti_join`

```{r anti_join}
anti_join(songs, artists, by = "name")
```

### per capita CO<sub>2</sub> emissions 

You'll join the [population](http://data.okfn.org/data/core/population) dataset to calculate per capita CO<sub>2</sub> emissions.

_**Task**. Summarize the total emissions by country  (not World or EU28) per capita across years from your long format table and return the top 5 emitting countries for **2014**._

_**Task**. Summarize the total emissions by country (not World or EU28) per capita across years from your long format table and return the top 5 emitting countries on **average** across all years._ 

```{r}
library(dplyr)
library(readxl)
library(readr)
library(tidyr)
library(ggplot2)
require('gridExtra')

popn = read_csv(
  'https://raw.githubusercontent.com/datasets/population/master/data/population.csv')
names(popn)[c(1,4)]=c("Country","Population")

co2e = read_csv('./data/co2e.csv')
co2elong=gather(co2e,key=Country)
names(co2elong)=c("Country","Year","CO2e")
co2elong$Year=as.numeric(co2elong$Year)

co2eANDpopn=left_join(popn,co2elong)
co2eANDpopn=co2eANDpopn[co2eANDpopn$Country!="World"&co2eANDpopn$Country!="EU28",]

co2eANDpopn[["CO2ePerCap"]]=co2eANDpopn[["CO2e"]]/co2eANDpopn[["Population"]]

emissions2014=co2eANDpopn[co2eANDpopn$Year==2014,]
top5emittersPC=emissions2014[order(emissions2014$CO2ePerCap,decreasing=TRUE),][1:5,]
top5emittersPC
top5emittersTOT=emissions2014[order(emissions2014$CO2e,decreasing=TRUE),][1:5,]
top5emittersTOT

AvgEmissions=group_by(co2eANDpopn,Country)%>%
  summarize(mean(CO2e,na.rm=TRUE))%>%
  na.omit()
AvgEmissions

group_by(co2eANDpopn,Country)%>%
  summarize(mean(CO2ePerCap,na.rm=TRUE))%>%
  na.omit()

top5emittersAVG=AvgEmissions[order(AvgEmissions[[2]],decreasing=TRUE),][1:5,]
top5emittersAVG

AvgYearlyEmissionsPC=group_by(co2eANDpopn,Year)%>%
  summarize(mean(CO2ePerCap,na.rm=TRUE))%>%
  na.omit()

AvgYearlyEmissionsTOT=group_by(co2eANDpopn,Year)%>%
  summarize(mean(CO2e,na.rm=TRUE))%>%
  na.omit()

names(AvgYearlyEmissionsPC)=c("Year","PerCapCO2e")
names(AvgYearlyEmissionsTOT)=c("Year","AvgCO2e")

AvgYearlyEmissions=left_join(AvgYearlyEmissionsPC,AvgYearlyEmissionsTOT)

#Created some graphics on the group's data as well

examplegraph1=ggplot(AvgYearlyEmissions,aes_string(x="Year",y="AvgCO2e"))+geom_line()+xlab(NULL)
examplegraph2=ggplot(AvgYearlyEmissions,aes_string(x="Year",y="PerCapCO2e"))+geom_line()+xlab(NULL)
grid.arrange(examplegraph1, examplegraph2, ncol=2)

```
